---
title: "Movie Industry Project"
author: "By Jesse"
output: html_document
---

##### In this I project I explore movie industry data and look for correlation between variables.

##### There are 6820 movies in the dataset (220 movies per year, 1986-2016). This data was scraped from IMDb.

#### Each movie has the following attributes:

* budget: the budget of a movie. Some movies don't have this, so it appears as 0
* company: the production company
* country: country of origin
* director: the director
* genre: main genre of the movie.
* gross: revenue of the movie
* name: name of the movie
* rating: rating of the movie (R, PG, etc.)
* released: release date (YYYY-MM-DD)
* runtime: duration of the movie
* score: IMDb user rating
* votes: number of user votes
* star: main actor/actress
* writer: writer of the movie
* year: year of release

<br>

Let's get started!

Loading in packages, removing objects from the environment and setting the working directory

```{r packages and wd, message=FALSE, warning=FALSE}
if (!require("pacman")) install.packages("pacman"); library(pacman)
pacman::p_load(tidyverse, caTools, lubridate,
               ggthemes, janitor, corrplot, GGally, psych)

rm(list=ls())
setwd("/Users/Jesse/R/MovieIndustryProject/")
```

Reading in the data

```{r reading in the data, message=FALSE, warning=FALSE}
df <- read_csv("/Users/Jesse/R/MovieIndustryProject/movies.csv")
```

Let's take a took at the data

```{r checking out the data}
head(df)
str(df)
```

Finally, let's check for any missing values:

```{r missing}
any(is.na(df))
```

No missing data! Let's move on to basic data cleaning.

```{r data cleaning}
df$year_corrected <- as.numeric(str_sub(df$released, 1, 4)) # fixing years not matching

df <- df %>% arrange(desc(gross)) # sorting by highest grossing movies
options(scipen = 999) # get rid of scientific notation

get_dupes(df)

head(df)
```

We created a new column to fix an issue with the release date and year not matching. Then we sort by highest grossing movies and check for any duplicates.

<br>

#### Looking for correlation

```{r corr}
df %>% ggplot(aes(budget, gross)) + geom_point(color = "red", alpha = .5) + 
    ggtitle("Budget vs Gross Earnings",) + theme_bw() + theme(plot.title = element_text(hjust = .5)) +
    geom_smooth(method = "lm")
```
<br>

Above is a scatterplot showing the relationship between budget and gross earnings.

```{r corr2}
df_cor <- df[,c(1,6,10,11,13,16)] # using only numeric features
cor(df_cor)
ggpairs(df_cor) # matrix of plots

cor_mx <- cor(df_cor) 
corrplot(cor_mx,method = "color", title = "Correlation Matrix for Numeric Features", 
         mar=c(0,0,1,0))
```
<br>

It looks like budget and votes have fairly strong positive correlation with gross earnings.

Next, let's convert the character columns to factors then numeric values to be included in the correlation matrix

```{r corr3}
# create numeric factor for all character columns
numeric_factor <- function(x){
    as.numeric(as.factor(x))
}

df_numfac <- df # creating frame to change characters to factored numeric values
df_numfac[,c(2:5,7,8,12,14)] <- lapply(df_numfac[,c(2:5,7,8,12,14)], numeric_factor) # converting non-numeric
                                                                                     # to factored numeric

cor(df_numfac[,c(-9,-15)]) # correlation with all features - removing old release date and release year
cor_numfac_mx <- cor(df_numfac[,c(-9,-15)]) # creating correlation matrix 
corrplot(cor_numfac_mx, method = "color", title = "Correlation Matrix for All Features", 
         mar=c(0,0,1,0)) # correlation plot
```
<br>

It looks like both budget and votes are still the most highly correlated with gross earnings, both showing fairly strong positive correlation.

Finally, let's look only at gross earnings in the correlation matrix

```{r corr4}
cor_gross <- cor(df_numfac[,c(-6,-9,-15)], df_numfac$gross) # matrix of only correlation w/ gross
cor_gross <- as.matrix(cor_gross[order(cor_gross[,1], decreasing = T),]) # ordered by highest correlation
cor_gross

colnames(cor_gross) <- c("gross") # changing column name

high_cor_gross <- cor_gross[cor_gross[,1] >= .5,, drop = F] # budget and votes have highest correlation w/ gross
high_cor_gross
```
<br>

There you have it. Budget and votes have the highest correlation with gross earnings. This is about what you would expect, having a high budget for a movie generally produces a better product, thus earning more. It was interesting to learn the correlation between user votes and the gross earnings for a movie. I was somewhat surprised company did not have a higher correlation with gross earnings. I would have expected some companies, such as Disney, 20th Century Fox, Universal Pictures, etc. to have bring in more money based on their brand name. 






